\name: A hierarchical Bayesian model for Supernova Cosmology

### 2018-04-09 papers/methods/paper.tex
Responding to more of Bobs feedback

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{tikz}
\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
\usetikzlibrary{bayesnet}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Almost two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}. Since that time, the number of observed Type Ia supernovae (SN Ia) have increased by more than an order of magnitude thanks to modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, Sako2014}. Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014} have been combined with complimentary probes of large scale structureand the CMB. For a recent review, see \citet{Huterer2018}. Despite these prodigious efforts, the nature of dark energy remains an unsolved mystery.

@@

One final benefit of ABC methods is that they can move past the traditional treatment of supernovae with summary statistics ($m_B$, $x_1$ and $c$). \citet{Jennings2016} presents both a metric used to compare forward modelled summary statistic populations (denoted the `Tripp' metric) and a metric directly applicable to the observed supernova light curves themselves, although evaluation of systematic uncertainty was only performed using the Tripp metric.

@@

Sitting comfortably between the traditional models simplicity and the complexity of forward modelling lies hierarchical Bayesian Models. With the introduction of multiple layers in our model, we can add far more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood. \citet{Mandel2009}, \citet{Mandel2011a} and \citet{Mandel2017} construct a hierarchical model that they apply to supernova light curve fitting. \citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide increased flexibility with reduced uncertainty over the traditional method, but does not incorporate bias correction. \citet{March2014, Karpenka2015} improve upon this by incorporating redshift-dependent magnitude corrections from \citet{Perrett2010} to remove bias, and validate on 100 realisations of SNLS like simulations. The recent BAHAMAS model \citep{Shariff2016} builds on this and reanalyses the JLA dataset (using their redshift dependent bias corrections), whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$. \citet{Ma2016} performed a reanalysis of the JLA dataset within a Bayesian formulation, finding significant differences in $\alpha$ and $\beta$ values from the original analysis from \citet{Betoule2014}. Notably, these methods rely on data that is bias corrected or ignore biases, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}. The simplification made by the UNITY analysis is that the bias is well described by an analytic function, and this function is fixed such that selection effects are assumed to be determined with perfect precision. Their model was validated on thirty realisations of statistics.

@@

We construct our hierarchical Bayesian model with several goals in mind: creation of a redshift-dependent underlying supernova population, increased treatment of systematics, and analytic correction of selection effects, including systematic uncertainty on those corrections. As this is closest to the UNITY method from \citet[][hereafter denoted \rubin]{Rubin2015}, we follow a similar model scaffold, and construct the model in Stan \citep{Carpenter2017, StanDevelopmentTeam2017} which uses automatic differentiation and the no-U-turn Sampler (NUTS), which is a variant of Hamiltonian Monte Carlo, to efficiently sample high dimensional parameter space.

At the most fundamental level, a supernova analysis is simply a mapping from an underlying population onto an observed population, where cosmology is encoded directly in the mapping function. The difficulty arises both in adequately describing the biases in the mapping function, and in adding sufficient, physically motivated flexibility in both of these populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model.

@@

The first layer of the hierarchy represents the parameters that describe each supernova. For the moment, let us consider a single supernova, and represent its parameters as $\theta$. The underlying supernova model and intrinsic dispersion model gives the probability of $p(y|\theta)$ for obtaining light curve data $y$. As mentioned earlier, we work with summary statistics -- numbers output from a light curve fitter given the input $y$. The function $f$ describes the action of the fitter $\lbrace \hat{m}_B, \hat{x}_1., \hat{c} \rbrace = f(y)$. For notational simplify, let us denote $\eta \equiv \lbrace m_B, x_1, c \rbrace$. The model prediction for our derived data is

@@

where $\mathcal{C}$ is a covariance matrix reported from light curve fitting. However, the presence of chromatic smearing in the intrinsic supernova population \citep{Guy2010, Chotard2011} means the summary statistics reported from light curve fitting are not Gaussian and equation \eqref{eq:pop} is violated. We address this problem with a similar method to previous analytic likelihoods -- including extra dispersion in the reported covariance matrix $\mathcal{C}$.

We model the extra dispersion only in colour, and do so by adding extra independent uncertainty on the colour observation. Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ was also tested, but it showed negligible improvement over just colour dispersion and was far more computationally inefficient. As shown in \citep{Kessler2013}, the extra colour dispersion shows heavy redshift dependence. As the extra dispersion may arise from underlying supernova physics, rather than Malmquist bias, we decide to incorporate redshift dependence in our extra uncertainty. We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature), where $\kappa_0$ and $\kappa_1$ are subject to strong Cauchy priors centered on zero and with width $0.002$, to enforce minimum possible extra uncertainty in our model. The value of this prior was determined from running many iterations with differing prior strengths on our simple data set with simulated extra dispersion according to the results of \citet{Kessler2013}. Each survey has its own set of $\kappa$ parameters. As such, our combined covariance on the observation $\hat{\eta}$ is given by $\mathcal{C}_{\rm tot} = \mathcal{C} + {\rm DiagMatrix}\left[0, 0, (\kappa_0 + \kappa_1 z)^2\right]$.

Furthermore, we use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift, and subtract this shift from our observed data. We allow the amount of shift to be parametrised to account for different intrinsic scatter models, where parameter $s_c$ controls the fraction of bias subtraction -- 0 for no shift, up to 1 for the maximum shift found in simulations \citep[which corresponds to the intrinsic scatter model from][]{Chotard2011}. Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a universal shift in supernovae colour. We place a uniform prior on $s_c$.

As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$. Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable. One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, those can naturally be modelled and taken into account.

@@

We follow prior work and model the colour population as an independent redshift-dependent skew normal for each survey. For the stretch population, we adopt a redshift-dependent normal, and magnitude dispersion is modelled as a normal. Following {\rubin} we allow the mean colour and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes. These nodes are represented as $\langle x_1^i \rangle$ and $\langle c^i \rangle$. Both the colour and stretch means are modelled with normal priors. Initial versions of our model adopted a fully covariant multivariate skew normal (with skewness set to zero only for the magnitude component), however pathological fitting complications required us to simplify our treatment. We parameterise the standard skew normal skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$. The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors, however are sampled in log space for efficiency in sampling close to zero.

As such, the only constant between survey populations is the absolute magnitude $M_B$, with the skewness, redshift-dependent means and width fit individually for each survey. The probability for a supernova to have true values $M_B$, $x_1$, $c$ given the underlying population is thus given as

@@

On top of the Ia populations, as described above, we also include a simplistic outlier population that also follows {\rubin} \citep[and therefore ][]{Kunz2007} as a Gaussian mixture; where the mean of the population is fixed to the SN Ia population, but the population width is set to a width of $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$. With the spectroscopic DES sample, the contamination rate is expected to be far too low to actually fit contamination population, however in future works with photometric samples that will suffer from significantly more contamination, it will be required that extra degrees of freedom are afforded the outlier population. Proof of concept analysis on simple simulation data show that an acceptable parametrisation is to represent the typically brighter contaminant population as $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle - \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations. For the purposes of the DES spectroscopic sample, which will be confirmed Type Ia supernovae, $\delta_{M_B}^{\rm outl} =0$. We assume that supernovae fall into the SN Ia population as determined by their observed classification probability $\hat{p}$, which is set to unity for the spectroscopic sample.

@@

It is now well known that host galaxy environment has a significant effect on supernova properties. The latest sample of over 1300 spectroscopically confirmed Type Ia supernovae show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}. The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014} invoke a step function such that $\Delta M = 0.08 \mathcal{H}(\log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function and $M$ is the galaxy mass in solar masses. The scale of this step function varies from analysis to analysis, with the 0.08 value shown previously sourced from \cite{Sullivan2010} and used in \citet{Betoule2014}. In this work we adopt the model used in {\rubin}, which follows the work from \citet{Rigault2013}, such that we introduce two parameters to incorporate a redshift-dependent host galaxy mass correction:

\begin{equation}
\Delta M = \delta(0) \left[ \frac{1.9\left(1 - \frac{\delta(0)}{\delta(\infty)}\right)  }{0.9 + 10^{0.95z}} + \frac{\delta(0)}{\delta(\infty)}\right].
\end{equation}

We also take flat priors on the parametrisation $\delta(0)$, $\delta(0)/\delta(\infty)$. With this correction, our calculation of absolute magnitude becomes

\begin{align}
M_B = m_B - \mu(z) - \alpha x_1 + \beta c - \Delta M \times m. \label{eq:l2}
\end{align}

@@

where the denominator represents an integral over all potential data. As $\theta$ represents the vector of all model parameters, and $D$ represents a vector of all observed variables, this is not a trivial integral. Techniques to approximate this integral, such as Monte-Carlo integration or high-dimensional Gaussian processes failed to give tractable posterior surfaces that could be sampled efficiently by HMC, and post-fitting importance sampling failed due to high-dimensionality (a brief dismissal of many months of struggle). We therefore simplify the integral and approximate the selection effects from their full expression in all of $\theta$-space, to apparent magnitude and redshift space independently (not dependent on $x_1$ or $c$), such that the denominator of equation \eqref{eq:main}, denoted now $w$ for simplicity, is given as

\begin{align}
w &= \int  \left[ \int P(S|m_B) P(m_B | z, \theta)\, d m_B \right] P(S|z) P(z|\theta)\, dz. \label{eq:w1}
\end{align}

@@

where $\Phi^c$ the complimentary CDF and $\mu_{\rm CDF}$ and $\sigma_{\rm CDF}$ specify the selection function. The appropriateness of an error function has been found by many past surveys \citep{Dilday2008, Barbary2010, Perrett2012, Graur2013, Rodney2014}. However, for surveys which suffer from saturation and thus rejection of low-redshift supernovae, or for groups of surveys treated together (as is common to do with low-redshift surveys), we find that a skew normal is a suitable analytic form, taking the form

@@

The selection functions are fit to apparent magnitude efficiency ratios calculated from SNANA simulations, by calculating an efficiency ratio as a function of apparent magnitude. Uncertainty of the Malmquist bias from the imperfection of the analytic function is incorporated into the fit. Uncertainty was uniformly added in quadrature (similar to adding intrinsic dispersion as a fit parameter) to the efficiency ratio data from our simulations until the reduced $\chi^2$ of the analytic fit reached $1$, allowing us to extract an uncertainty covariance matrix for our analytic fits to either the error function or the skew normal.

With the well sampled approximation we can remove the redshift integral in Eq \eqref{eq:w1} and replace it with a correction for each observed supernova. For the error function (denoted with the subscript `CDF') and skew normal selection functions respectively (denoted with a subscript `Skew'), this correction becomes

\begin{align}
w_{\rm CDF} &= \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
w_{\rm Skew} &= 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
&\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellowz}
\end{align}

and is incorporated into our likelihood.

@@

We generate simple toy data to validate the basic premise of the model. For both high-redshift and low-redshift (LowZ) data we draw from an underlying $M_B$, $x_1$, $c$ population and translate into apparent magnitude space using $\Omega_m = 0.3$, $\alpha= 0.14$ and $\beta = 3.1$. Mass probabilities are randomly drawn from the interval 0 to 1, and a mass correction with $\delta(0) = 0.08$ and $\delta(0)/\delta(\infty) =0.5$ included. Absolute magnitudes are drawn from $\mathcal{N}(-19.3, 0.1)$, true stretch values are drawn from $\mathcal{N}(0, 1)$ and true colour values are drawn from $\mathcal{N}^{\rm Skew}(0, 0.1, 2)$. Redshifts are drawn from a power law to increase the number of events as redshift increases.

>>>>

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{tikz}
\usepackage{enumerate}
\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
\usetikzlibrary{bayesnet}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Almost two decades have passed since the discovery of the accelerating universe \citep{Riess1998, Perlmutter1999}. Since that time, the number of observed Type Ia supernovae (SNIa) have increased by more than an order of magnitude thanks to modern surveys at both low redshift \citep{Bailey2008, Freedman2009, Hicken2009,  Contreras2010, Conley2011}, and higher redshift \citep{Astier2006, Wood-Vasey2007, Frieman2008, Balland2009, Amanullah2010, Sako2014}. Cosmological analyses of these supernova samples \citep{Kowalski2008, Kessler2009, Conley2011, Suzuki2012, Betoule2014, Rest2014} have been combined with complimentary probes of large scale structureand the CMB. For a recent review, see \citet{Huterer2018}. Despite these prodigious efforts, the nature of dark energy remains an unsolved mystery.

@@

One final benefit of ABC methods is that they can move past the traditional treatment of supernovae with summary statistics ($m_B$, $x_1$ and $c$). \citet{Jennings2016} presents both two metrics (which are used to measure the distance between the forward modelled population and observed population, which is minimised in fitting). The first metric compares forward modelled summary statistic populations (denoted the `Tripp' metric) and the second utilises the observed supernova light curves themselves, moving past summary statistics. However, we note that evaluation of systematic uncertainty was only performed using the Tripp metric.

@@

Sitting between the traditional models simplicity and the complexity of forward modelling lies Bayesian hierarchical models (BHM). Hierarchical models utilise multiple layers of connected parameters in their models, with the layers linked via well defined and physical motivated conditional probabilities. For example, an observation of a parameter from a population will be conditioned on the true value of the parameter, which itself will be conditioned on the population distribution of that parameter. We can thus easily incorporate different distributions, populations, and parameter inter-dependence which cannot be found in traditional analyses where uncertainty must be encapsulated in a covariance matrix.

With the introduction of multiple layers in our model, we can add far more flexibility than a traditional analysis whilst still maintaining most of the computational benefits that come from having a tractable likelihood. \citet{Mandel2009, Mandel2011a, Mandel2017} construct a hierarchical model that they apply to supernova light curve fitting. \citet{March2011} derive a hierarchical model and simplify it by analytically marginalising over nuisance parameters to provide increased flexibility with reduced uncertainty over the traditional method, but does not incorporate bias correction. \citet{March2014, Karpenka2015} improve upon this by incorporating redshift-dependent magnitude corrections from \citet{Perrett2010} to remove bias, and validate on 100 realisations of SNLS like simulations. The recent BAHAMAS model \citep{Shariff2016} builds on this and reanalyses the JLA dataset (using their redshift dependent bias corrections), whilst including extra freedom in the correction factors $\alpha$ and $\beta$, finding evidence for redshift dependence on $\beta$. \citet{Ma2016} performed a reanalysis of the JLA dataset within a Bayesian formulation, finding significant differences in $\alpha$ and $\beta$ values from the original analysis from \citet{Betoule2014}. Notably, these methods rely on data that is bias corrected or ignore biases, however the UNITY framework given by \citet{Rubin2015} incorporates selection efficiency analytically in the model, and is applied to the Union 2.1 dataset \citep{Suzuki2012}. The simplification made by the UNITY analysis is that the bias is well described by an analytic function, and this function is fixed such that selection effects are assumed to be determined with perfect precision. Their model was validated on thirty realisations of statistics.

@@

We construct our hierarchical Bayesian model with several goals in mind: creation of a redshift-dependent underlying supernova population, increased treatment of systematics, and analytic correction of selection effects, including systematic uncertainty on those corrections. As this is closest to the UNITY method from \citet[][hereafter denoted \rubin]{Rubin2015}, we follow a similar model scaffold, and construct the model in Stan \citep{Carpenter2017, StanDevelopmentTeam2017}. The primary challenge of fitting hierarchical models is their large number of fit parameters, and Stan, which uses automatic differentiation and the no-U-turn Sampler (NUTS) (a variant of Hamiltonian Monte Carlo), allows us to efficiently sample high dimensional parameter space.

At the most fundamental level, a supernova analysis is simply a mapping from an underlying population onto an observed population, where cosmology is encoded directly in the mapping function. The difficulty arises both in adequately describing the biases in the mapping function, and in adding sufficient, physically motivated flexibility in both of these populations whilst not adding \textit{too} much flexibility, such that model fitting becomes pathological due to increasing parameter degeneracies within the model. In the following sections, we will describe these layers and mapping functions.

@@

As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$. Potential sources of redshift error (such as peculiar velocities) is taken into account not via uncertainty on redshift (which is technically challenging to implement) but instead uncertainty on distance modulus. This is discussed further in Section \ref{sec:systreat}. Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable. One of the strengths of this model (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, those can naturally be modelled and taken into account.

The first layer of the hierarchy represents the parameters that describe each supernova. That is, observed parameters are denoted with a hat, whilst the true value is denoted without a hat. For example, $c$ is the true colour of the supernova, whilst $\hat{c}$ is the colour we observe, which, as it has uncertainty, is different from $c$.

For the moment, let us consider a single supernova, and represent its true parameters as $\theta$. The underlying supernova model and intrinsic dispersion model gives the probability of $p(y|\theta)$ for obtaining light curve data $y$. As mentioned earlier, we work with summary statistics -- numbers output from a light curve fitter given the input $y$. The function $f$ describes the action of the fitter $\lbrace \hat{m}_B, \hat{x}_1., \hat{c} \rbrace = f(y)$. For notational simplify, let us denote $\eta \equiv \lbrace m_B, x_1, c \rbrace$. The model prediction for our derived data is

@@

where $\mathcal{C}$ is a covariance matrix reported from light curve fitting. Unfortunately, this assumption of Gaussianity is incorrect. The intrinsic supernovae population is commonly described in terms of coherent magnitude dispersion, wherein some supernovae are uniformly brighter or dimmer than the population mean, and in chromatic dispersion, whereby the spectral energy density of the supernovae varies non-uniformly as a function of wavelength. The presence of chromatic smearing in the intrinsic supernova population \citep{Guy2010, Chotard2011} means the summary statistics reported from light curve fitting are under-reported, not Gaussian and equation \eqref{eq:pop} is violated. We address this problem with a similar method to previous analytic likelihoods: including extra dispersion in the reported covariance matrix $\mathcal{C}$. Whilst this doesn't solve the issue of non-Gaussianity, it does address the under-reporting of uncertainty, which is the primary source of bias.

We model the extra dispersion only in colour, and do so by adding extra independent uncertainty on the colour observation. Fully covariant extra dispersion on $\lbrace m_B, x_1, c \rbrace$ was also tested, but it showed negligible improvement over just colour dispersion and was far more computationally inefficient. As shown in \citep{Kessler2013}, the extra colour dispersion shows heavy redshift dependence. As the extra dispersion may arise from underlying supernova physics, rather than Malmquist bias, we decide to incorporate redshift dependence in our extra uncertainty. We thus add $\kappa_0 + \kappa_1 z$ to our observed colour uncertainty (in quadrature), where $\kappa_0$ and $\kappa_1$ are subject to strong Cauchy priors centered on zero and with width $0.002$, to enforce minimum possible extra uncertainty in our model. The $\kappa$ parameters are tightly constrained as they are highly degenerate with the width of the intrinsic colour population $\sigma_c$, and the specific value of this prior was determined from running many iterations of simulation fitting with differing prior strengths on our simple data set with simulated extra dispersion according to the results of \citet{Kessler2013}. The value $0.002$ can be varied by a factor of approximately three in either direction whilst giving stable results, with larger values (more than a factor of three) for the prior resulted in incorrect determination of $\sigma_c$, whilst smaller values resulted in being unable to capture the extra dispersion  Each survey has its own set of $\kappa$ parameters. As such, our combined covariance on the observation $\hat{\eta}$ is given by $\mathcal{C}_{\rm tot} = \mathcal{C} + {\rm DiagMatrix}\left[0, 0, (\kappa_0 + \kappa_1 z)^2\right]$.

Furthermore, we use fiducial simulations (discussed further in Section \ref{sec:selection}) to characterise the shift in the mean colour measured as a function of redshift, and subtract this shift from our observed data. With the extra dispersion introduced from the $\kappa$ parameters addressing the under-reporting of error, shifting the mean colour address the non-Gaussianity introduced by chromatic smearing. We allow the amount of shift to be parametrised to account for different intrinsic scatter models, where parameter $s_c$ controls the fraction of bias subtraction -- 0 for no shift, up to 1 for the maximum shift found in simulations \citep[which corresponds to the intrinsic scatter model from][]{Chotard2011}. Unlike $\kappa_0$ and $\kappa_1$, which are redefined for each survey, $s_c$ represents a universal shift in supernovae colour. We place a uniform prior on $s_c$. It is important to note that the effects being modelled and characterised by simulations here are features of the supernova population, not cosmological features, and so the choice of fiducial cosmology has negligible impact.

@@

We follow prior work and model the colour population as an independent redshift-dependent skew normal for each survey. For the stretch population, we adopt a redshift-dependent normal, and magnitude dispersion is modelled as a normal. Following {\rubin} we allow the mean colour and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes. These nodes are represented as $\langle x_1^i \rangle$ and $\langle c^i \rangle$. Both the colour and stretch means are modelled with normal priors. Initial versions of our model adopted a fully covariant multivariate skew normal (with skewness set to zero only for the magnitude component), however pathological fitting complications required us to simplify our treatment. We parameterise the skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$, which allows $\alpha_c$ to span positive values in the realm of physical plausibility as determined from constraints in \citet{Scolnic2016}. The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors, however are sampled in log space for efficiency in sampling close to zero.

As such, the only constant between survey populations is the mean absolute magnitude $\langle M_B \rangle$, with the skewness, redshift-dependent means and width fit individually for each survey. The probability for a supernova to have true values $M_B$, $x_1$, $c$ given the underlying population is thus given as

@@

On top of the SNIa populations, as described above, we also include a simplistic outlier population that also follows {\rubin} \citep[and therefore ][]{Kunz2007} as a Gaussian mixture; where the mean of the population is fixed to the SNIa population, but the population width is set to a width of $\sigma^{\rm outl} = 1$ in $M_B$, $x_1$ and $c$. With the spectroscopic DES sample, the contamination rate is expected to be far too low to actually fit contamination population, however in future works with photometric samples that will suffer from significantly more contamination, it will be required that extra degrees of freedom are afforded the outlier population. This contamination is not just a result of non-SNIa being observed, but can also arise from host galaxy misidentification and incorrect redshifts - depending on the cuts on the optimised ratio between purity and efficiency, this can result in between 3\% to 9\% misidentification  \citep{Gupta2016}. Proof of concept analysis on simple simulation data show that an acceptable parametrisation to deal with contamination is to represent the typically dimmer contaminant population as $\langle M_B^{\rm outl} \rangle = \langle M_B \rangle + \delta_{M_B}^{\rm outl}$, where $\delta_{M_B}^{\rm outl}$ is constrained to be positive, or even to be greater than a small positive number to reduce degeneracy between the two populations. For the purposes of the DES spectroscopic sample, which will be confirmed SNIa, $\delta_{M_B}^{\rm outl} =0$. We assume that supernovae fall into the SNIa population as determined by their observed classification probability $\hat{p}$, which is set to unity for the spectroscopic sample.

@@

It is now well known that host galaxy environment has a significant effect on supernova properties. The latest sample of over 1300 spectroscopically confirmed Type Ia supernovae show $>5\sigma$ evidence for correlation between host mass and luminosity \citep{Uddin2017}. The traditional correction, as employed in analyses such as \citet{Suzuki2012} and \citet{Betoule2014} invoke a step function such that $\Delta M = 0.08 \mathcal{H}(\log(M) - 10))$, where $\mathcal{H}$ is the Heaviside step function, $M$ is the galaxy mass in solar masses and $0.08$ represents the size of the magnitude step. The scale of this step function varies from analysis to analysis, with the 0.08 value shown previously sourced from \cite{Sullivan2010} and used in \citet{Betoule2014}. In this work we adopt the model used in {\rubin}, which follows the work from \citet{Rigault2013}, such that we introduce two parameters to incorporate a redshift-dependent host galaxy mass correction:

\begin{equation}
\Delta M = \delta(0) \left[ \frac{1.9\left(1 - \frac{\delta(0)}{\delta(\infty)}\right)  }{0.9 + 10^{0.95z}} + \frac{\delta(0)}{\delta(\infty)}\right], \label{eq:mass}
\end{equation}

where $\delta(0)$ represents the correction at redshift zero, and $\delta(\infty)$ a parameter allowing the behaviour to change with increasing redshift. We take flat priors on the parametrisation $\delta(0)$, $\delta(0)/\delta(\infty)$. With this correction, our calculation of absolute magnitude becomes

\begin{align}
M_B = m_B - \mu(z) - \alpha x_1 + \beta c - \Delta M \times {\rm mass\_prob}. \label{eq:l2}
\end{align}

@@

where the denominator represents an integral over all potential data $D$. As $\theta$ represents the vector of all model parameters, and $D$ represents a vector of all observed variables, this is not a trivial integral. Techniques to approximate this integral, such as Monte-Carlo integration or high-dimensional Gaussian processes failed to give tractable posterior surfaces that could be sampled efficiently by Hamiltonian Monte-Carlo, and post-fitting importance sampling failed due to high-dimensionality (a brief dismissal of many months of struggle). We therefore simplify the integral and approximate the selection effects from their full expression in all of $\theta$-space, to apparent magnitude and redshift space independently (not dependent on $x_1$ or $c$), such that the denominator of equation \eqref{eq:main}, denoted now $d$ for simplicity, is given as

\begin{align}
d &= \int  \left[ \int P(S|m_B) P(m_B | z, \theta)\, d m_B \right] P(S|z) P(z|\theta)\, dz. \label{eq:w1}
\end{align}

@@

where $\Phi^c$ the complimentary cumulative distribution function (CDF) and $\mu_{\rm CDF}$ and $\sigma_{\rm CDF}$ specify the selection function. The appropriateness of an error function has been found by many past surveys \citep{Dilday2008, Barbary2010, Perrett2012, Graur2013, Rodney2014}. However, for surveys which suffer from saturation and thus rejection of low-redshift supernovae, or for groups of surveys treated together (as is common to do with low-redshift surveys), we find that a skew normal is a suitable analytic form, taking the form

@@

The selection functions are fit to apparent magnitude efficiency ratios calculated from SNANA simulations, by calculating an efficiency ratio as a function of apparent magnitude. Uncertainty of the Malmquist bias from the imperfection of the analytic function is incorporated into the fit. Uncertainty was uniformly added in quadrature (similar to adding intrinsic dispersion as a fit parameter) to the efficiency ratio data from our simulations until the reduced $\chi^2$ of the analytic fit reached one, allowing us to extract an uncertainty covariance matrix for our analytic fits to either the error function or the skew normal.

With the well sampled approximation we can remove the redshift integral in Eq \eqref{eq:w1} and replace it with a correction for each observed supernova. For the error function (denoted with the subscript `CDF') and skew normal selection functions respectively (denoted with a subscript `Skew'), the correction \textit{per SNIa} becomes

\begin{align}
d_{\rm CDF/SNIa} &= \Phi^c\left(  \frac{m^*_B - \mu_{\rm CDF}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm CDF}^2}}  \right) \label{eq:seldes}\\
d_{\rm Skew} &= 2\mathcal{N}\left( \frac{m_B^* - \mu_{\rm Skew}}{\sqrt{\sigma^{*2}_{m_B} + \sigma_{\rm Skew}^2}}\right) \times \notag\\
&\quad\quad \Phi\left(  \frac{{\rm sign}(\alpha_{\rm Skew})(m_B^* - \mu_{\rm Skew})}{\frac{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}{\sigma^2_{\rm Skew}} \sqrt{\frac{\sigma^2_{\rm Skew}}{\alpha^2_{\rm Skew}} + \frac{\sigma_{m_B}^{*2} \sigma^2_{\rm Skew}}{\sigma_{m_B}^{*2} + \sigma^2_{\rm Skew}}} }\right), \label{eq:sellowz}
\end{align}

and is incorporated into our likelihood. This is illustrated in Figure \ref{fig:efficiency} for aid in visualisation.

\begin{figure}
	\begin{center}
		\includegraphics[width=\columnwidth]{efficiency.pdf}
	\end{center}
	\caption{The efficiency for supernova discovery at an arbitrary redshift. Shown in both panels in dashed blue is the SNIa population distribution, which takes the form of a normal distribution. The top panel shows a CDF based survey efficiency (green dotted line), whilst the bottom panel shows a skew normal based survey efficiency (red dotted line), but as generic functions of apparent magnitude. The survey efficiency, given the SNIa population, is shown as a solid line in both panels, and the probability of observing a SNIa is found by integrating over the population detection efficiency as described in equation \eqref{eq:w1}, and has been shown by shading the area integrated. This area is what is analytically given by equations \eqref{eq:seldes} and \eqref{eq:sellowz}.}
	\label{fig:efficiency}
\end{figure}

@@

We generate simple toy data to validate the basic premise of the model. Data generation algorithm is described below:

\begin{enumerate}[1.]
	\item Draw a redshift from a power law distribution. For the low redshift survey this is $\mathcal{U}(0.0004, 0.01)^{0.5}$, and for the DES-like survey this is $\mathcal{U}(0.008, 1.0)^{0.3}$.
	\item Draw a true absolute magnitude, stretch and colour from the respective distributions $\mathcal{N}(-19.3, 0.1)$, $\mathcal{N}(0, 1)$, $\mathcal{N}(0, 0.1)$.
	\item Draw a random mass probability from $\mathcal{U}(0, 1)$ and calculate the mass-brightness correction using $\delta(0) = 0.08$ and $\delta(0)/\delta(\infty) = 0.5$ and equation \eqref{eq:mass}.
	\item Calculate $\mu(z)$ given the drawn redshift and cosmological parameters $\Omega_m = 0.3$, $w = -1$. Use this to determine the true apparent magnitude of the object $m_B$ using equation \eqref{eq:l2}.
	\item Determine if the SNIa is detected using detection probability $P(S|m_B) = \mathcal{N}^{\rm skew}(13.72, 1.35, 5.87)$ for the low redshift survey (numeric values obtained by fitting to existing low redshift data). For the DES-like survey, accept with probability $P(S|m_B) = \Phi^C(23.14, 0.5)$. Repeat from step one until we have a supernova which passes.
	\item Add independent, Gaussian observational error onto the true $m_B$, $x_1$, $c$ using Gaussian widths of $0.04$, $0.2$, $0.03$ respectively (following the mean uncertainty for DES-like SNANA simulations). Add extra colour uncertainty in quadrature of $\kappa_0 + \kappa_1 z$, where $\kappa_0 = \kappa_1 = 0.03$.
\end{enumerate}

For both high-redshift and low-redshift (LowZ) data we draw from an underlying $M_B$, $x_1$, $c$ population and translate into apparent magnitude space using $\Omega_m = 0.3$, $\alpha= 0.14$ and $\beta = 3.1$. Mass probabilities are randomly drawn from the interval 0 to 1, and a mass correction with $\delta(0) = 0.08$ and $\delta(0)/\delta(\infty) =0.5$ included. Absolute magnitudes are drawn from $\mathcal{N}(-19.3, 0.1)$, true stretch values are drawn from $\mathcal{N}(0, 1)$ and true colour values are drawn from $\mathcal{N}^{\rm Skew}(0, 0.1, 2)$. Redshifts are drawn from a power law to increase the number of events as redshift increases.
